% id: order-connected
% title: Order Connected Sets are Measurable in \(\mathbb{R}^n\)
% date: 30 Jul. 2022

\documentclass{article}
\usepackage{amsmath, amssymb}

\begin{document}

Yaël Dillies was reading a book by Bollobas in which he claimed without proof that an order connected set (with respect to the pointwise ordering)
in \(\mathbb{R}^n\) is measurable. Yaël asked on Discord if anyone knows how to prove it. This short note contains my proof of this fact.

\begin{definition}
Given a partially ordered set \((X, (\le))\), \(S \subseteq X\) is said to be order connected if for all \(x, y \in S\),
\[[x, y] := \{z \mid x \le z \le y\} \subseteq S.\]
\end{definition}

Equipping \(\mathbb{R}^n\) with the point-wise ordering (i.e. for
\(x := (x_i)_{i = 1}^n,  y := (y_i)_{i = 1}^n \in \mathbb{R}^n\), \(x \le y\) if and only if
\(x_n \le y_n\) for all \(n = 1, \cdots, n\)), \(\mathbb{R}^n\) form a partially ordered set and we
can therefore talk about its order connected subsets.

\begin{theorem}
\(S \subseteq \mathbb{R}^n\) is Lebesgue measurable if \(S\) is order connected with respect to the
point-wise ordering on \(\mathbb{R}^n\).
\end{theorem}

\begin{proof}
Defining \(S^+ := \{x \mid \exists \ y \in S, y \le x\}\) and \(S^- := \{x \mid \exists \ y \in s, x \le y\}\), we
observe \(S = S^+ \cap S^-\). Therefore, it is sufficient to show that \(S^+\) and \(S^-\) are measurable.
We will show measurability for \(S^+\) while the case for \(S^-\) is similar.

It is clear that for all \(x \in \mathbb{R}^n\), \(\{x\}^+\) is measurable. So, defining \(Q := S^+ \cap \mathbb{Q}^n\),
\[Q^+ = \bigcup_{q \in Q} \{q\}^+ \subseteq S^+\]
is measurable. Furthermore, \(S^+ \setminus Q^+ \subseteq \partial S^+\) since for all \(x \in S^\circ\), there exists some
open neighborhood \(B \subseteq S^\circ\) of \(x\); so, as \(Q\) is dense in \(S^+\), there exists an element
\(q \in Q \cap B\) such that \(q \le x\) and hence, \(x \in \{q\}^+ \subseteq Q^+\). Now, as the Lebesgue \(\sigma\)-algebra is
complete and \(S^+ = Q^+ \cup (S^+ \setminus Q^+)\), it suffices to show that \(\partial S^+\) is a null-set.

To show \(\text{Leb}(\partial S^+) = 0\) we will invoke the Lebesgue density theorem. Namely, by showing
\[\partial S^+ \subseteq \{x \in \overline{S^+} \mid d(x) \notin \{0, 1\}\}\]
where
\[d(x) := \lim_{\epsilon \to 0} d_\epsilon(x) = \lim_{\epsilon \to 0}
    \frac{\text{Leb}(\overline{S^+} \cap B_\epsilon(x))}{\text{Leb}(B_\epsilon(x))}\]
as \(\overline{S^+}\) is closed and hence measurable, the Lebesgue density theorem tells us
\[\text{Leb}(\partial S^+) \le \text{Leb}(\{x \in \overline{S^+} \mid d(x) \notin \{0, 1\}\}) = 0.\]

Indeed, for all \(x \in \partial S^+, \epsilon > 0\), \(\overline{S^+} \cap B_\epsilon(x)\) must contain
the with the right-upper quadrant of the ball by the very definition of \(S^+\), implying
\(\text{Leb}(\overline{S^+} \cap B_\epsilon(x)) \gtrapprox 2^{-n} \text{Leb}(B_\epsilon(x))\) bounding
\(d(x)\) from below. Similarly, \(\overline{S^+}^c \cap B_\epsilon(x)\) must contain the left-bottom
quadrant of the ball as otherwise \(x\) is contained by some \(\{y\}^+\) for some \(y \in \overline{S}^+\)
contradicting \(x \in \partial S^+\). Hence, \(\text{Leb}(\overline{S^+} \cap B_\epsilon(x)) \lessapprox
(1 - 2^{-n}) \text{Leb}(B_\epsilon(x))\) bounding \(d(x)\) from above and so, proving the required
inclusion.

\end{proof}

\end{document}
